---
title: "Translating prior predictive distributions into priors for model parameters"
subtitle: "SURPH away day"
author: "Andrew Manderson"
date: "2022-05-27"
output:
  beamer_presentation:
    includes:
      in_header:
        tex-input/pre.tex
    keep_tex: yes
    theme: metropolis
    latex_engine: xelatex
    slide_level: 1
    incremental: no
fontsize: 14pt
# bibliography: bibliography/multi-phi-bib.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, fig.align = "center")
```

# An example

- You find yourself modelling human growth, measured in centimetres, for a uniformly random sample of adolescents

# Possible prior predictive

Possible prior predictive $\textcolor{myredhighlight}{t(Y)} =$

```{r ppd_one}
knitr::include_graphics("figures/pb-model/pop-target-marginal.pdf")
```

# Model (1)

- "_The Preece-Baines model is a standard model for this kind of data_"
- The simplest PB model looks like this:

\begin{equation*}
  Y(\texttt{age}) = h_{1} - \frac{
    2(h_{1} - h_{0})
  } {
    \exp\left\{s_{0}(\texttt{age} - \gamma)\right\}
    \exp\left\{s_{1}(\texttt{age} - \gamma)\right\}
  }
\end{equation*}

# Model (2)

\begin{equation*}
  Y(\texttt{age}) = h_{1} - \frac{
    2(h_{1} - h_{0})
  } {
    \exp\left\{s_{0}(\texttt{age} - \gamma)\right\}
    \exp\left\{s_{1}(\texttt{age} - \gamma)\right\}
  }
\end{equation*}

- How should you "translate" $\pd(Y)$ into a joint prior for $\theta = (h_{1}, h_{0}, s_{0}, s_{1}, \gamma)$?
- Best practice is _painful_ (especially for more complex models)

# General idea (1)

- Pick hyperparameters $\lambda$ of prior $\pd(\theta \mid \lambda)$ by minimising some discrepancy

\begin{equation*}
  D(\lambda) = \int d\left(
  \textcolor{blues2}{\pd(Y \mid \lambda)}, \textcolor{myredhighlight}{t(Y)}\right) \text{d}Y
\end{equation*}

- Solution $\lambda^{*} = \min_{\lambda} D(\lambda)$

# General idea (2)

- Solution $\textcolor{blues2}{\pd(Y \mid \lambda^{*})}$ and target $\textcolor{myredhighlight}{t(Y)}$:

```{r}
knitr::include_graphics("figures/pb-model/pop-target-marginal-with-samples.pdf")
```

# Challenges

- Optimisation surface
- Numerics
- Inherent, irreducible noise
- Underspecification

# Partial solutions

We can partly address these issues:

- Two-stage, multi-objective, gradient-free global optimisation
- Careful numerical implementation and importance sampling
- Regularisation term(s)
- `R` package implementation with few requirements for users: [\textcolor{blue}{gitlab.com/andrew-manderson/pbbo}](https://gitlab.com/andrew-manderson/pbbo)
